---
title: "Covid91 vaccine study FinalFall2022"
author: "Rebecca Castro"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males
```{r}
malesdata=filter(FinalData,sex=="M")
```

$h_0$: The treatment has no affect on males.

$h_A$: The treatment is a affect on males.

### Graphical Description
```{r}
barchartGC(~infected + treatment,data=malesdata)
barchartGC(~infected + treatment,data=malesdata, type="percent")

```

Based on the graph we see that twice the amount of males received placebo while infected with Covid921 that those with the treatment. With the graph it appears that the treatment is more effective for those infected and with treatment.

## Numerical Description

```{r}
table1 <- xtabs(~infected + treatment, data=malesdata)
rowPerc(table1)
colPerc(table1)
```
According to the numerical data, it reflects that 68.44% of males contracted covid921 received the placebo and 31.56% of males received the treatment. The data shows that the treatment is effective for males.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi squared test has a p-value of 0.00000000005307 which is less than 0.05 meaning we reject the null hypothesis. $h_A$: The treatment is a affect on males. According to the p-value we concluded that the treatment does have an affect on the males.
The fisher test has a p-value of 0.0000000000002364 which is less than 0.05 meaning we reject the null hypothesis. According to the p-value we conclude that the treatment does have an affect on the males.

## Discussion and Conclusion
Based on the data above, among the males that contract covid921 68.44% received the placebo and 31.65% received the treatment. According to the data, among the males the odd are that 2.18 times likely to contract covid921 than males who received the treatment.

## Females

```{r}
femalesdata=filter(FinalData,sex=="F")
```
$h_0$: The treatment has no affect on females.

$h_A$: The treatment is a affect on females.

### Graphical Description
```{r}
barchartGC(~infected + treatment,data=femalesdata)
barchartGC(~infected + treatment,data=femalesdata, type="percent")

```

Based on the graph we see that more females received treatment while infected with covid921 than those who received the placebo. the females are more effective while infected and treatment.

## Numerical Description

```{r}
table1 <- xtabs(~infected + treatment, data=femalesdata)
rowPerc(table1)
colPerc(table1)
```
According to the numerical data, it reflects that 42.82% of females contracted covid921 received the placebo and 57.18% of females received the treatment. The data shows that the treatment is effective for females.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi squared test has a p-value of 0.005 which is equal to 0.05 meaning we accept the null hypothesis. According to the p-value we concluded that the treatment doesn't have an affect on the females.
The fisher test has a p-value of 0.004 which is less than 0.05 meaning we reject the null hypothesis. According to the p-value we conclude that the treatment does have an affect on the females.

## Discussion and Conclusion
Based on the data above, among the females that contract covid921 42.82% received the placebo and 57.18% received the treatment. 

## LGBTQ

```{r}
lgbtqdata=filter(FinalData,sex=="LGBTQ")
```
$h_0$: The treatment has no affect on LGBTQ.

$h_A$: The treatment is a affect on LGBTQ.


### Graphical Description
```{r}
barchartGC(~infected + treatment,data=lgbtqdata)
barchartGC(~infected + treatment,data=lgbtqdata, type="percent")

```
according to the graph having being infected or treatment does not matter in the lgbtq.

## Numerical Description

```{r}
table1 <- xtabs(~infected + treatment, data=lgbtqdata)
rowPerc(table1)
colPerc(table1)
```
### Inferential Results




## Druggies
```{r}
druguserdata=filter(FinalData,sex=="DrugUser")
```

### Graphical Description
```{r}
barchartGC(~infected + treatment,data=druguserdata)
barchartGC(~infected + treatment,data=druguserdata, type="percent")

```

## Numerical Description
```{r}
table1 <- xtabs(~infected + treatment, data=druguserdata)
rowPerc(table1)
colPerc(table1)
```
### Inferential Results

# Overall Results and Conclusions

The overall results show that the sex is affect by covid and not if you use drugs or in the lgbtq.